---
title: "DataConstruct"
author: "Susanna"
date: "2026-09-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

# Construct social network

Taken from FIS data manual:

rm(list=ls())
packages <- c("sna","tidyverse","haven")
lapply(packages, require, character.only=T)
rm(packages)
path_in <- "PATH_TO_DATA"

fis.orig <- read_dta(paste(path_in,"/fis_long.dta",sep=""))
fis.comp <- read_dta(paste(path_in,"/fis_composition.dta",sep=""))

 # Specify Set of Network Characteristics (10 friendship nominations)
netw <- c("a1_0101","a1_0102","a1_0103","a1_0104","a1_0105",
"a1_0106","a1_0107","a1_0108","a1_0109","a1_0110")
 # Specify Covariate for Node Color: ed_a6 (gender)
cova <- c("ed_a6")
 # Specify Individual Identifier
iden <- c("id_p")
 # Specify Grade and Wave
g <- 1
w <- 1

comp.g.w <- filter(fis.comp,wave == w, id_g == g)
fis.g.w <- filter(fis.orig,wave == w, id_g == g) %>%
select(iden,netw,cova) %>%
mutate(right = 1)
fis.g.w <- left_join(comp.g.w,fis.g.w,by="id_p") %>% arrange(id_p)

fis.fr.g.w <- select(fis.g.w,iden,netw)
ids <- unlist(select(fis.fr.g.w,"id_p")) # extract all IDs
34
ids.s <- sort(ids) # sorts unique IDs
ids.tr <- cbind(ids.s,1:length(ids)) # generates matching IDs from 1 onwards

miss.w <- ids[unlist(is.na(select(fis.g.w,right)))]

f <- 0
for (i in ids.s) {
f <- f + 1
fis.fr.g.w[fis.fr.g.w == i] <- f
}
miss.w.tr <- ids.tr[,2][ids.s %in% miss.w]

edge.w <- cbind(unlist(fis.fr.g.w[, 1]), unlist(c(fis.fr.g.w[, -1]))) # generate edgelist
mat.w <- matrix(0,nrow=length(ids),ncol=length(ids)) # generate empty adjacency matrix
mat.w[edge.w] <- 1 # fill adjacency matrix
mat.w[miss.w.tr,] <- NA # fill missing actors with NAs
mat.adj.w <- as.network(mat.w,matrix.type="adjacency") # transform to R adjacency matrix

# Creating Identification Variable
Children who reported having no parent or grandparent born outside of Germany were instructed to skip the section on heritage identification. I therefore firstly need to make a distinction between children with no non-German ancestry and those who do have non-German ancestry. To do so, I make use of the generational status variable, coded by the data collectors based on the procedure set out by Dollmann et al. (2014). Only students with no non-German ancestry are coded as native. All others, regardless of generational status, are coded as having migrant ancestry. Cases of German-born children where information on parents or grandparents was missing were coded as incomplete. Cases where all information was missing were coded as missings (NA).

Subsequently, within the migrant ancestry group, children were coded based on their identification as described in the Data section. Those who self-identified as only German and reported being seen as only German by Germans were included in the majority group, as their migrant ancestry in this case is unlikely to be relevant for their social situation. 
This may for example apply to children with distant non-German ancestry (for example, who only have one grandparent born outside of Germany).

Incomplete cases who did not respond to the identification question were assumed to be native-born following the instruction to skip this question. In other cases, they were treated the same way as children in the migrant ancestry group.

For full missings on the ancestry data, identification was leading, with perceived other-categorisation by Germans used to differentiate migrantised German students from the majority group.

```{r}
#no parent/grandparent born abroad --> majority
#if gen status is not in dataset: df$fammig <- ifelse(b7_coded_group==1 & b8_coded_group==1 & b9_1==1 & b9_2==1 & b9_3==1 & b9_4==1, "Maj", "Mig")

#df$fammig <- df$ed_miggen_cils_narrow
#df$fammig <- recode_values(fammig,
                           #1 ~ "Mig",
                          # 2 ~ "Mig",
                          # 3 ~ "Mig",
                          # 4 ~ "Mig",
                          # 5 ~ "Mig",
                          # 6 ~ "Mig",
                          # 7 ~ "Mig",
                          # 8 ~ "Mig",
                          # 9 ~ "Mig",
                          # 10 ~ "Mig",
                          # 11 ~ "Mig",
                          # 12 ~ "Mig",
                          # 13 ~ "Mig",
                          # 14 ~ "Maj",
                          # 15 ~ "Mig",
                          # 16 ~ "Mig",
                          # 17 ~ "Mig",
                          # 18 ~ "Mig",
                          # 19 ~ "Incomplete",
                          # 20 ~ "Incomplete",
                          # 21 ~ NA)

#creating identity groups, strict requirement for dual identity
#df$identity <- df$fammig
#df %>% mutate(identity = replace_when(identity, 
        #identity == "Mig" & c4==3  ~ "3",
        #identity == "Mig" & (c4==4 | c4==5) ~ "4",
        #identity == "Mig" & (c4==1 | c4==2) & c5_1 >= 2 ~ "2",
        #identity == "Mig" & (c4==1 | c4==2) & (c5_1==1 | c5_1==. | c5_1==.u) ~ "1",
        #identity == "Mig" & (c4==. | c4==.u) ~ NA,
        #identity == "Incomplete" & (c4==. | c4==.u) & (c5_1==. | c5_1==.u) ~ "1",
        #identity == "Incomplete" & (c4==. | c4==.u) & c5_1 !=. & c5_1 != .u ~ NA,
        #identity == "Incomplete" & (c4==1 | c4==2) & c5_1 >= 2 ~ "2",
        #identity == "Incomplete" & (c4==1 | c4==2) & (c5_1==1 | c5_1==. | c5_1==.u) ~ "1",
        #identity == "Incomplete" & c4==3 ~ "3",
        #identity == "Incomplete" & (c4==4 | c4==5) ~ "4",
        #identity == NA & (c4==1 | c4==2) & (c5_1 == 1 | c5_1==. | c5_1==.u) ~ "1",
        #identity == NA & (c4==1 | c4==2) & c5_1 >= 2 ~ "2",
        #identity == NA & c4==3 ~ "3",
        #identity == NA & (c4==4 | c4==5) ~ "4"))
#levels(df$identity)
#summary(df$identity)

#1 is majority, 2 is migrantised, 3 is dual, 4 is heritage

#df$dual <- ifelse(identity == NA, NA, ifelse(identity == 3, 1, 0))
#levels(df$dual)
#summary(df$dual)

#1 is dual identifier, 0 is mono-identifier
```

For robustness check with lenient dual identity definition:

```{r}
#df$identityrob <- df$fammig
#df %>% mutate(identityrob = replace_when(identityrob, 
        #identityrob == "Mig" & (c4==2 | c4==3 | c4==4) ~ "3",
        #identityrob == "Mig" & (c4==5) ~ "4",
        #identityrob == "Mig" & (c4==1) & c5_1 >= 2 ~ "2",
        #identityrob == "Mig" & (c4==1) & (c5_1==1 | c5_1==. | c5_1==.u) ~ "1",
        #identityrob == "Mig" & (c4==. | c4==.u) ~ NA,
        #identityrob == "Incomplete" & (c4==. | c4==.u) & (c5_1==. | c5_1==.u) ~ "1",
        #identityrob == "Incomplete" & (c4==. | c4==.u) & c5_1 !=. & c5_1 != .u ~ NA,
        #identityrob == "Incomplete" & c4==1 & c5_1 >= 2 ~ "2",
        #identityrob == "Incomplete" & c4==1 & (c5_1==1 | c5_1==. | c5_1==.u) ~ "1",
        #identityrob == "Incomplete" & (c4==2 | c4==3 | c4==4) ~ "3",
        #identityrob == "Incomplete" & c4==5 ~ "4",
        #identityrob == NA & c4==1 & (c5_1 == 1 | c5_1==. | c5_1==.u) ~ "1",
        #identityrob == NA & c4==1 & c5_1 >= 2 ~ "2",
        #identityrob == NA & (c4==2 | c4==3 | c4==4) ~ "3",
        #identityrob == NA & c4==5 ~ "4"))
#levels(df$identityrob)
#summary(df$identityrob)

#1 is majority, 2 is migrantised, 3 is dual, 4 is heritage

#df$dualrob <- ifelse(identityrob == NA, NA, ifelse(identityrob == 3, 1, 0))
#levels(df$dualrob)
#summary(df$dualrob)

#1 is dual identifier, 0 is mono-identifier
```

