---
title: "Week3"
author: "Susanna"
date: "2026-09-14"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

## Swimming Lessons

Copy-paste the following:

```{r}
rm(list = ls())

fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

colorize <- function(x, color) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
}
```

Downloads uninstalled packages, saves everything based on date.

More packages:

```{r}
packages = c("RSiena", "devtools", "igraph")
fpackage.check(packages)
#devtools::install_github('JochemTolsma/RsienaTwoStep', build_vignettes=TRUE)
packages = c("RsienaTwoStep")
fpackage.check(packages)
```

```{r}
library(igraph)
net1g <- graph_from_adjacency_matrix(ts_net1, mode = "directed")
coords <- layout_(net1g, nicely())  #let us keep the layout
par(mar = c(0.1, 0.1, 0.1, 0.1))
{
    plot.igraph(net1g, layout = coords)
    graphics::box()
}
```

```{r}
set.seed(24553253)
ego <- ts_select(net = ts_net1)
ego
```

What are the options:

```{r}
options <- ts_alternatives_ministep(net = ts_net1, ego = ego)
options
plots <- lapply(options, graph_from_adjacency_matrix, mode = "directed")
par(mar = c(0, 0, 0, 0) + 0.1)
par(mfrow = c(2, 2))

fplot <- function(x) {
    plot.igraph(x, layout = coords, margin = 0)
    graphics::box()
}

lapply(plots, fplot)
```

How to choose between options? Let's say I like extra ties. Option with
more ties is evaluated more positively. Opt for most positively
evaluated option. If multiple maxima, choose randomly. But: this is
deterministic (I know everything and always make the right choice), not
stochastic (there is some randomness). So instead, we weigh options
based on evaluation.

```{r}
sapply(options, ts_degree, ego = ego) #degree for all options
sapply(options, ts_recip, ego = ego) #number of reciprocated ties
```

I decide degree and reciprocity are important statistics
(characteristics of ego or ego network, or of alter). Model will try
different parameters to see if they will result in a network similar to
the observed network at t2. You add up the benefits and downsides on all
effects for each option, which results in a certain likelihood(?) of
this option being chosen.

```{r}
option <- 4
ts_degree(options[[option]], ego = ego) * -1 + ts_recip(options[[option]], ego = ego) * 1.5
```

The above is the agent's evaluation of option 4 (based on degree and
reciprocity).

Doing that for all networks:

```{r}
eval <- sapply(options, FUN = ts_eval, ego = ego, statistics = list(ts_degree, ts_recip), parameters = c(-1,
    1.5))
eval
print("network with maximum evaluation score:")
which.max(eval)
```

This shows that option 4 is evaluated most positively, and therefore
most likely to be chosen (but stochastic not deterministic, so other
options are still possible - just less likely) We weigh likelihood based
on McFadden's choice function. All options need to have a chance between
0 and 1, sum of all chances should be 1. Probability of choosing i = exp
of evaluation score, divided by the sum of the chances of each of the
alternatives

```{r}
choice <- sample(1:length(eval), size = 1, prob = exp(eval)/sum(exp(eval)))
print("choice:")
choice
# print('network:') options[[choice]]
```

Repeat very very often. Choice of next agent is dependent on choice of
previous agent. Common stopping rule approach: add it as a parameter
(try if we get there with 2 tie changes per actor, or with 3, etc)

**Interpretation**

More strongly positive parameter: actors on average find variable more
important in making decision (so interpretation is on micro level!)
Negative parameter: actors try to avoid this Hypothesis should be about
out-degrees, not in-degrees Usually one tie change influences lots of
different things simultaneously (degree, but also for example
reciprocity)

For end project: show that you can interpret effects. Start simple:
positive or negative, then compare with other effects. Usually degree
parameter is negative, because chance of there being a tie is lower than
chance of no tie (density in network is lower than 0.5)

## Diving into the Deep

```{r}
fpackage.check <- function(packages) {
    lapply(packages, FUN = function(x) {
        if (!require(x, character.only = TRUE)) {
            install.packages(x, dependencies = TRUE)
            library(x, character.only = TRUE)
        }
    })
}

fsave <- function(x, file = NULL, location = "./data/processed/") {
    ifelse(!dir.exists("data"), dir.create("data"), FALSE)
    ifelse(!dir.exists("data/processed"), dir.create("data/processed"), FALSE)
    if (is.null(file))
        file = deparse(substitute(x))
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, ".rda", sep = "")
    save(x, file = totalname)  #need to fix if file is reloaded as input name, not as x. 
}

fload <- function(filename) {
    load(filename)
    get(ls()[ls() != "filename"])
}

fshowdf <- function(x, ...) {
    knitr::kable(x, digits = 2, "html", ...) %>%
        kableExtra::kable_styling(bootstrap_options = c("striped", "hover")) %>%
        kableExtra::scroll_box(width = "100%", height = "300px")
}

# this is the most important one. We created it in the previous script

f_pubnets <- function(df_scholars = df, list_publications = publications, discip = "sociology", affiliation = "RU",
    waves = list(wave1 = c(2018, 2019, 2020), wave2 = c(2021, 2022, 2023))) {

    publications <- list_publications %>%
        bind_rows() %>%
        distinct(title, .keep_all = TRUE)

    df_scholars %>%
        filter(affil1 == affiliation | affil2 == affiliation) %>%
        filter(discipline == discip) -> df_sel

    networklist <- list()
    for (wave in 1:length(waves)) {
        networklist[[wave]] <- matrix(0, nrow = nrow(df_sel), ncol = nrow(df_sel))
    }

    publicationlist <- list()
    for (wave in 1:length(waves)) {
        publicationlist[[wave]] <- publications %>%
            filter(gs_id %in% df_sel$gs_id) %>%
            filter(year %in% waves[[wave]]) %>%
            select(author) %>%
            lapply(str_split, pattern = ",")
    }

    publicationlist2 <- list()
    for (wave in 1:length(waves)) {
        publicationlist2[[wave]] <- publicationlist[[wave]]$author %>%
            # lowercase
        lapply(tolower) %>%
            # Removing diacritics
        lapply(stri_trans_general, id = "latin-ascii") %>%
            # only last name
        lapply(word, start = -1, sep = " ") %>%
            # only last last name
        lapply(word, start = -1, sep = "-")
    }

    for (wave in 1:length(waves)) {
        # let us remove all publications with only one author
        remove <- which(sapply(publicationlist2[[wave]], FUN = function(x) length(x) == 1) == TRUE)
        publicationlist2[[wave]] <- publicationlist2[[wave]][-remove]
    }

    for (wave in 1:length(waves)) {
        pubs <- publicationlist2[[wave]]
        for (ego in 1:nrow(df_sel)) {
            # which ego?
            lastname_ego <- df_sel$lastname[ego]
            # for all publications
            for (pub in 1:length(pubs)) {
                # only continue if ego is author of pub
                if (lastname_ego %in% pubs[[pub]]) {
                  aut_pot <- which.max(pubs[[pub]] %in% lastname_ego)
                  # only continue if ego is first author of pub
                  if (aut_pot == 1) {
                    # check all alters/co-authors
                    for (alter in 1:nrow(df_sel)) {
                      # which alter
                      lastname_alter <- df_sel$lastname[alter]
                      if (lastname_alter %in% pubs[[pub]]) {
                        networklist[[wave]][ego, alter] <- networklist[[wave]][ego, alter] + 1
                      }
                    }
                  }
                }
            }
        }
    }
    return(list(df = df_sel, network = networklist))
}
packages = c("RSiena", "tidyverse", "stringdist", "stringi")

fpackage.check(packages)

```

First figure out where your network is!

```{r}
df <- fload("./data/20230621df_complete.rda")
publications <- fload("./data/20230621list_publications_jt.rda")

output <- f_pubnets()
df_soc <- output[[1]]
df_network <- output[[2]]

view(df_soc)
view(df_network)
```

Separate the waves:

```{r}
wave1 <- df_network[[1]]
wave2 <- df_network[[2]]
# let us put the diagonal to zero
diag(wave1) <- 0
diag(wave2) <- 0
# we want a binary tie (not a weighted tie)
wave1[wave1 > 1] <- 1
wave2[wave2 > 1] <- 1
# put the nets in an array
net_soc_array <- array(data = c(wave1, wave2), dim = c(dim(wave1), 2)) #2 is the number of waves
# dependent
net <- sienaDependent(net_soc_array)
```

Adding another variable

```{r}
# gender
gender <- as.numeric(df_soc$gender == "female") #1 if f, 0 if m
gender <- coCovar(gender) #constant covariate
mydata <- sienaDataCreate(net, gender)
```

Now we have our **data object**!

Time for the effects:

```{r}
myeff <- getEffects(mydata)
effectsDocumentation(myeff)
```

```{r}
ifelse(!dir.exists("results"), dir.create("results"), FALSE)
print01Report(mydata, modelname = "./results/soc_init")
```

```{r}
myeff <- includeEffects(myeff, isolateNet, inPop, outAct, inAct, transTrip)  #adding the effects of interest
```

```{r}
myeff <- includeEffects(myeff, sameX, egoX, altX, interaction1 = "gender") #we're specifying that x means gender. Nothing to do with interaction effects!
```

```{r}
myAlgorithm <- sienaAlgorithmCreate(projname = "soc_init")
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff))
# (the outer parentheses lead to printing the obtained result on the screen) if necessary, estimate
# further
(ans <- siena07(myAlgorithm, data = mydata, effects = myeff, prevAns = ans, returnDeps = TRUE)) #estimating something based on algorithm on my data
```

The rate parameter is the average amount of mini-steps taken per agent,
which is 4.0795 in my case. Outdegree effect is quite strongly negative,
so we have a network with a low density. Reciprocity has a very strongly
positive effect. Transitive triplets also seem to be quite attractive.
Network isolate effect is also present (people who do not have any ties
are unlikely to make or receive new ones).

## Drowning???

Let's try to add a new effect I guess?

```{r}
myeff <- includeEffects(myeff, sharedPop)
myAlgorithm2 <- sienaAlgorithmCreate(projname = "soc_init")
(ans <- siena07(myAlgorithm2, data = mydata, effects = myeff))

(ans <- siena07(myAlgorithm2, data = mydata, effects = myeff, prevAns = ans, returnDeps = TRUE)) #estimating something based on algorithm on my data
```
Yay I managed to add the shared popularity effect! It's not significant :(. 

*Okay, now I'd like to try the whole fit thing.*

First we should define the functions:

```{r}
GeodesicDistribution <- function(i, data, sims, period, groupName, varName, levls = c(1:5, Inf), cumulative = TRUE,
    ...) {
    x <- networkExtraction(i, data, sims, period, groupName, varName)
    require(sna)
    a <- sna::geodist(symmetrize(x))$gdist
    if (cumulative) {
        gdi <- sapply(levls, function(i) {
            sum(a <= i)
        })
    } else {
        gdi <- sapply(levls, function(i) {
            sum(a == i)
        })
    }
    names(gdi) <- as.character(levls)
    gdi
}

testall <- function(ans) {
    for (i in which(ans$test)) {
        sct <- score.Test(ans, i)
        cat(ans$requestedEffects$effectName[i], "\n")
        print(sct)
    }
    invisible(score.Test(ans))
}
```

See if it works:

```{r}
gofi0 <- sienaGOF(ans, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
plot(gofi0)
```

Okay that worked! Can I do it with something other than the in-degree distribution?

First I'll try to calculate segregation (which is not working AAAAAAHHH)

```{r}
segregation <- sna::nacf(df_network, gender, type = "moran", neighborhood.type = "out", demean = TRUE)
segregation
```


```{r}
gofi0 <- sienaGOF(ans, IndegreeDistribution, verbose = FALSE, join = TRUE, varName = "net")
```

To do for next time: look at RSiena manual and read chapters, look at relative influence, update RQs and list of required data
