---
title: "Week4"
author: "Susanna"
date: "2026-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

### Playing with Plots

```{r}
require(igraph)
g <- make_graph("Zachary")
plot(g)
```

Apparently, there's a random component to how graphs are made: placement of nodes based on popularity.

```{r}
gmat <- as_adjacency_matrix(g, type = "both", sparse = FALSE)
gmat
```

g is in igraph format, we just turned it back into a matrix

```{r}
isSymmetric(gmat)
```

If symmetric, network is undirected

```{r}
# be aware that directed graphs are considered as undirected. but g is undirected.
igraph::transitivity(g, type = c("localundirected"), isolates = c("NaN", "zero"))
```
We're counting triangles now. No person is standing out.

```{r}
igraph::betweenness(g, directed = FALSE)
```
1 and 34 have a VERY high betweenness score. Not very visible in initial graph, so if we want to focus on this finding, we want a graph where 1 and 34 do not overlap so much with other nodes. 3 is also interesting: bridge beween bridge?

*How can we emphasise this in visualisation?* 

hue palet, vibrancy based on betweenness (standardised)
size based on betweenness
or change layout

```{r}
igraph::dyad_census(g)
```
```{r}
igraph::triad_census(g)
# I will use sna because it shows the names of the triads as well.
sna::triad.census(gmat)
unloadNamespace("sna")  #I will detach this package again, otherwise it will interfere with all kind of functions from igraph, and my students will hate me for that.
```

Transitivity: how many triads are closed divided by the total amount of triads that are or could have been closed.
300 is fully closed:
- last digit: number of dyads without ties
- second digit: number of dyads with one tie
- first digit: number of dyads with two ties (mutual dyads)
- specific subtype:
  C: cyclic
  D: downward
  U: upward
  T: transitive

Uitgaand van actorperspectief wordt één triade 3 keer geteld (één keer vanuit elke betrokken actor).

Manually calculating global transitivity:

```{r}
igraph::transitivity(g, type = "global")
sna::gtrans(gmat)

triad_g <- data.frame(sna::triad.census(gmat))

transitivity_g <- (3 * triad_g$X300)/(triad_g$X201 + 3 * triad_g$X300)
transitivity_g
```

(local transitivity would be transitivity in network of ego)

1 and 34 are not connected, both have high betweenness, but low local transitivity - let's visualise that story:

```{r}
# changing V
V(g)$size = betweenness(g, normalized = T, directed = FALSE) * 60 + 10  #after some trial and error
plot(g, mode = "undirected")
```

V is about vertices, e is about edges

To make it look nice in article, define size of plot window beforehand for alllll images! 

```{r}
set.seed(2345)
l <- layout_with_mds(g)  #https://igraph.org/r/doc/layout_with_mds.html
plot(g, layout = l)
```
```{r}
l  #let us take a look at the coordinates
l[1, 1] <- 4
l[34, 1] <- -3.5
plot(g, layout = l)
```
Second image suggests conflict, group being pulled apart by 34 and 1. 
In article: explain that 1 and 34 were manually placed at opposite sides of the network to illustrate the described process.

# Plotting On Ashwin's data

```{r}
load("./data/IndividualCharacteristics.RData")
load("./data/NetworkCharacteristics.RData")
```

```{r}
# view(studying[[1]])
studying_df_t1 <- as.data.frame(studying[[1]][1])
class(studying_df_t1)
```
```{r}
# view(weak_netw_t1)

studying_df_t2 <- as.data.frame(studying[[1]][2])
class(studying_df_t2)
```
```{r}
# view(studying_df_t2)
weak_netw_t2 <- data.matrix(studying_df_t2)
weak_netw_t2[weak_netw_t2==9]<-NA
class(weak_netw_t2)
```
```{r}
# view(hanging[[1]])
hanging_df_t1 <- as.data.frame(hanging[[1]][1])
class(hanging_df_t1)
```
```{r}
# view(hanging_df_t1)
strong_netw_t1 <- data.matrix(hanging_df_t1)
strong_netw_t1[strong_netw_t1==9]<-NA
class(strong_netw_t1)
```
```{r}
# view(strong_netw_t1)

hanging_df_t2 <- as.data.frame(hanging[[1]][2])
class(hanging_df_t2)
```
```{r}
# view(hanging_df_t2)
strong_netw_t2 <- data.matrix(hanging_df_t2)
strong_netw_t2[strong_netw_t2==9]<-NA
class(strong_netw_t2)
```
Making own matrix

```{r}
testm <- matrix(NA, nrow = 4, ncol = 4)
testm[1,1] <- 0
testm[1,2] <- 0
testm[1,3] <- NA
testm[3,1] <- 1
testm[3,2] <- 1
testm[4,2] <- 0
testm[4,2] <- 1
testm
mrow <- rowSums(is.na(testm)) == nrow(testm)
mcol <- colSums(is.na(testm)) == ncol(testm)
mtot <- which(mrow & mcol)
mtot
testm[which(mrow), ] <- 0
testm
```


```{r}
require(igraph)
uni1nona <- strong_netw_t1
uni1nona[is.na(uni1nona)] <- 0
uni1 <- igraph::graph_from_adjacency_matrix(uni1nona, mode = "directed", weighted = NULL, diag = TRUE, add.colnames = NA,
    add.rownames = NA)

plot(uni1)
```
Yay I have a graph!!!

```{r}
V(uni1)$size = igraph::degree(uni1) * 1.05 + 5
plot(uni1)
```
```{r}
plot(uni1, 
     edge.arrow.size = 0.1,
     vertex.label.cex = 0.4,
     vertex.size = 10)
```
```{r}
V(uni1)$size = igraph::degree(uni1)*0.5 + 5
plot(uni1, 
     edge.arrow.size = 0.1,
     vertex.label.cex = 0.4)
```

```{r}
w1international <- international[[1]]
w1international <- w1international[, "W1INTERNATIONAL"]
w1international
print(uni1)

#edges <- igraph::as_data_frame(uni1, what = "edges")
#edges
#uni2 <- graph_from_data_frame(edges, directed = TRUE, vertices = studentids)
#vertex_attr_names(uni1)

#V(uni1)$international <- w1international

#V(uni1)$color <- ifelse(V(uni1)$w1international == 1, "blue", "orange")
length(w1international)
length(V(uni1))


```
I think I lost the connection to the original IDs. Let's try again

```{r}
class(hanging_df_t2)
hangingt2nona <- hanging_df_t2
hangingt2nona[hangingt2nona==9]<- 0
hangingt2nona <- as.matrix(hangingt2nona)


hangingt2 <- graph_from_adjacency_matrix(hangingt2nona)
plot(hangingt2)

```
```{r}
V(hangingt2)$size = igraph::degree(hangingt2)*0.5 + 5
plot(hangingt2, 
     edge.arrow.size = 0.1,
     vertex.label.cex = 0.4)
```
```{r}
#check nina's lab journal
```

# Webscraping tutorial

```{r}
#install.packages("xml2")
require(xml2)
#install.packages("rvest")
require(rvest)
```
```{r}
url <- "https://nos.nl/zoeken?q=meertaligheid&page=1"

nos <- read_html(url)


#elements: alle elementen, element: één element
lijst <- nos %>% html_elements("main") %>%
  html_elements("ul") %>%
  html_elements("li")

lijst[1] %>% html_children %>% html_attr(name="href")

lijst[1] %>% html_elements("time") %>% 
  html_attr(name="datetime")

```

