---
title: "Week4"
author: "Susanna"
date: "2026-09-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

# Playing with Plots

```{r}
require(igraph)
g <- make_graph("Zachary")
plot(g)
```

Apparently, there's a random component to how graphs are made: placement of nodes based on popularity.

```{r}
gmat <- as_adjacency_matrix(g, type = "both", sparse = FALSE)
gmat
```

g is in igraph format, we just turned it back into a matrix

```{r}
isSymmetric(gmat)
```

If symmetric, network is undirected

```{r}
# be aware that directed graphs are considered as undirected. but g is undirected.
igraph::transitivity(g, type = c("localundirected"), isolates = c("NaN", "zero"))
```
We're counting triangles now. No person is standing out.

```{r}
igraph::betweenness(g, directed = FALSE)
```
1 and 34 have a VERY high betweenness score. Not very visible in initial graph, so if we want to focus on this finding, we want a graph where 1 and 34 do not overlap so much with other nodes. 3 is also interesting: bridge beween bridge?

*How can we emphasise this in visualisation?* 

hue palet, vibrancy based on betweenness (standardised)
size based on betweenness
or change layout

```{r}
igraph::dyad_census(g)
```
```{r}
igraph::triad_census(g)
# I will use sna because it shows the names of the triads as well.
sna::triad.census(gmat)
unloadNamespace("sna")  #I will detach this package again, otherwise it will interfere with all kind of functions from igraph, and my students will hate me for that.
```

Transitivity: how many triads are closed divided by the total amount of triads that are or could have been closed.
300 is fully closed:
- last digit: number of dyads without ties
- second digit: number of dyads with one tie
- first digit: number of dyads with two ties (mutual dyads)
- specific subtype:
  C: cyclic
  D: downward
  U: upward
  T: transitive

Uitgaand van actorperspectief wordt één triade 3 keer geteld (één keer vanuit elke betrokken actor).

Manually calculating global transitivity:

```{r}
igraph::transitivity(g, type = "global")
sna::gtrans(gmat)

triad_g <- data.frame(sna::triad.census(gmat))

transitivity_g <- (3 * triad_g$X300)/(triad_g$X201 + 3 * triad_g$X300)
transitivity_g
```

(local transitivity would be transitivity in network of ego)

1 and 34 are not connected, both have high betweenness, but low local transitivity - let's visualise that story:

```{r}
# changing V
V(g)$size = betweenness(g, normalized = T, directed = FALSE) * 60 + 10  #after some trial and error
plot(g, mode = "undirected")
```

V is about vertices, e is about edges

To make it look nice in article, define size of plot window beforehand for alllll images! 

```{r}
set.seed(2345)
l <- layout_with_mds(g)  #https://igraph.org/r/doc/layout_with_mds.html
plot(g, layout = l)
```
```{r}
l  #let us take a look at the coordinates
l[1, 1] <- 4
l[34, 1] <- -3.5
plot(g, layout = l)
```
Second image suggests conflict, group being pulled apart by 34 and 1. 
In article: explain that 1 and 34 were manually placed at opposite sides of the network to illustrate the described process.

# Plotting On Ashwin's data

```{r}
load("./data/IndividualCharacteristics.RData")
load("./data/NetworkCharacteristics.RData")
```

```{r}
# view(hanging[[1]])
hanging_df_t1 <- as.data.frame(hanging[[1]][1])
class(hanging_df_t1)
```
```{r}
# view(hanging_df_t1)
strong_netw_t1 <- data.matrix(hanging_df_t1)
strong_netw_t1[strong_netw_t1==9]<-NA
class(strong_netw_t1)
```
```{r}
# view(strong_netw_t1)

hanging_df_t2 <- as.data.frame(hanging[[1]][2])
class(hanging_df_t2)
```
```{r}
# view(hanging_df_t2)
strong_netw_t2 <- data.matrix(hanging_df_t2)
strong_netw_t2[strong_netw_t2==9]<-NA
class(strong_netw_t2)
```
Making own matrix

```{r}
testm <- matrix(NA, nrow = 4, ncol = 4)
testm[1,1] <- 0
testm[1,2] <- 0
testm[1,3] <- NA
testm[3,1] <- 1
testm[3,2] <- 1
testm[4,2] <- 0
testm[4,2] <- 1
testm
mrow <- rowSums(is.na(testm)) == nrow(testm)
mcol <- colSums(is.na(testm)) == ncol(testm)
mtot <- which(mrow & mcol)
mtot
testm[which(mrow), ] <- 0
testm
```


```{r}
require(igraph)
uni1nona <- strong_netw_t1
uni1nona[is.na(uni1nona)] <- 0
uni1 <- igraph::graph_from_adjacency_matrix(uni1nona, mode = "directed", weighted = NULL, diag = TRUE, add.colnames = NA,
    add.rownames = NA)

plot(uni1)
```
Yay I have a graph!!!

```{r}
V(uni1)$size = igraph::degree(uni1) * 1.05 + 5
plot(uni1)
```
```{r}
plot(uni1, 
     edge.arrow.size = 0.1,
     vertex.label.cex = 0.4,
     vertex.size = 10)
```
```{r}
V(uni1)$size = igraph::degree(uni1)*0.5 + 5
plot(uni1, 
     edge.arrow.size = 0.1,
     vertex.label.cex = 0.4)
```

```{r}
w1international <- international[[1]]
w1international <- w1international[, "W1INTERNATIONAL"]
w1international
print(uni1)

#edges <- igraph::as_data_frame(uni1, what = "edges")
#edges
#uni2 <- graph_from_data_frame(edges, directed = TRUE, vertices = studentids)
#vertex_attr_names(uni1)

#V(uni1)$international <- w1international

#V(uni1)$color <- ifelse(V(uni1)$w1international == 1, "blue", "orange")
length(w1international)
length(V(uni1))


```
I think I lost the connection to the original IDs. Let's try again

```{r}
class(hanging_df_t2)
hangingt2nona <- hanging_df_t2
hangingt2nona[hangingt2nona==9]<- 0
hangingt2nona <- as.matrix(hangingt2nona)


hangingt2 <- graph_from_adjacency_matrix(hangingt2nona, mode = "directed", weighted = NULL, diag = TRUE, add.colnames = NA, add.rownames = NA)
plot(hangingt2)

```
Ugly version succeeded. Now, change size based on degree and make it a little prettier:
```{r}
V(hangingt2)$size = igraph::degree(hangingt2)*0.5 + 5
plot(hangingt2, 
     edge.arrow.size = 0.1,
     vertex.label.cex = 0.4)
```
Not bad!
Okay this is what failed last time, let's see if Nina's code saves me:
```{r}
#first isolate international wave 2
w2international <- international[[1]]
w2international <- w2international[, "W2INTERNATIONAL"]
w2international

#credits to nina's lab journal
colourinternational <- ifelse(w2international==1, "purple", "cyan4") #I thought I tried this before???
plot(hangingt2,
     vertex.color = colourinternational, # nice color for the vertices
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     vertex.label.dist = 0.9,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)

```
Yay it worked! It automatically makes all NAs white - that's nice! I only see one international, and they're very isolated... :( 

Let's try the same with a more equally distributed attribute like gender:
```{r}
#first isolate gender wave 2
w2gender <- gender[[1]]
w2gender <- w2gender[, "T2GENDER"]
w2gender

#colour based on gender
colourgender <- ifelse(w2gender==1, "thistle1", "aquamarine")

#plot
plot(hangingt2,
     vertex.color = colourgender, # nice color for the vertices
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     vertex.label.dist = 0.9,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```
Oooooh fun! Interesting, it seems like there isn't a lot of gender segregation going on. I guess people have somewhat grown out of that by the time they start uni!

I do feel like the size is a bit too small now. Let's see if we can do something about that:

```{r}
colourgender2 <- ifelse(w2gender==1, "thistle1", "aquamarine")
V(hangingt2)$size = igraph::degree(hangingt2)*0.7 + 7
plot(hangingt2,
     vertex.color = colourgender, # nice color for the vertices
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     vertex.label.dist = 0.5,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```

Trying some other things in graphic design of plot:
```{r}
V(hangingt2)$size = igraph::degree(hangingt2)*0.7 + 7
plot(hangingt2,
     vertex.color = colourgender, # nice color for the vertices
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     vertex.label.dist = 0.5,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1,
     edge.color = "grey75")
```

I like this! I feel like I'm seeing three clusters, two bigger ones and a smaller one. Particularly in the two bigger groups, there's people with a high degree. Towards the edge of the graph, degrees are smaller. 39 does not have a very high degree, but does seem to form a bridge between the two bigger groups. 
Maybe I can do something with betweenness?

```{r}
V(hangingt2)$size = igraph::betweenness(hangingt2)*0.7 + 7
plot(hangingt2,
     vertex.color = colourgender, # nice color for the vertices
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     vertex.label.dist = 0.9,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```

OOF. Looks like some people have such a high betweenness score that this does not work at all. Maybe if I normalise it though?

```{r}
V(hangingt2)$size = betweenness(hangingt2, normalized = T, directed = TRUE) * 60 + 7
plot(hangingt2,
     vertex.color = colourgender, # nice color for the vertices
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     vertex.label.dist = 0.9,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```

Hmmm weird. We don't actually see huge differences now. I kind of want to take a look at the betweenness scores themselves...

```{r}
igraph::betweenness(hangingt2, directed = TRUE)
```
I do see a lot of variation in here. Not sure why that is not turning up in the graph. Maybe the normalised scores are very small. I could multiply with a higher number:

```{r}
V(hangingt2)$size = betweenness(hangingt2, normalized = T, directed = TRUE) * 250 + 7
plot(hangingt2,
     vertex.color = colourgender, # nice color for the vertices
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     vertex.label.dist = 0.9,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```

That's better! We still see the two big groups, with members who have a high betweenness score. 13, 9, and 18 stand out with their middle position between the two groups. 39's betweenness score is actually not as large as I thought it would be. Good to know!

Now let's try transitivity, which may be relevant to understand the closeness of the observed friend/'hanging' groups.

```{r}
V(hangingt2)$size = igraph::transitivity(hangingt2, type = "local", isolates = "zero")*10 + 7
plot(hangingt2,
     vertex.color = colourgender, # nice color for the vertices
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     vertex.label.dist = 0.5,  # we'll pull the labels a bit away from the vertices
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1)
```

Cool! The triads at the edges of the graph have a relatively high transitivity score. 83 and 95, who had a very high degree, don't stand out now. What is their local transitivity score actually?
```{r}
igraph::transitivity(hangingt2, type = "local", isolates = "zero")
```
Yeah, their transitivity is very average. So we have two people (83 and 95) who are hanging out with lots of others, who sort of cluster around them. Their friends however are not necessarily super connected to one another, so it doesn't seem like they are super tight friend groups - just two popular kids. 83 and 95 are not hanging out together. 

I think the one based on degree is most informative of this story. So let's get that one back on the screen again and see if we can make it even prettier:

```{r}
colourgender2 <- ifelse(w2gender==1, "goldenrod1", "firebrick")
V(hangingt2)$size = igraph::degree(hangingt2)*0.9 + 7
E(hangingt2)$width <- edge_betweenness(hangingt2) * 0.01 # Edge width based on betweenness
plot(hangingt2,
     vertex.color = adjustcolor(colourgender2, alpha.f = .4),
     vertex.frame.color = "gray",  # we'll put a gray frame around vertices
     vertex.label.color = "black",  # not that ugly blue color for the labels (names)
     vertex.label.cex = 0.4,  # make the label a bit smaller too
     edge.curved = 0.2, # curved edges is always a nice touch
     edge.arrow.size = 0.1,
     edge.color = "gray",
     layout=layout.fruchterman.reingold, main="fruchterman.reingold") #based on attraction between connected notes and repulsion between non-connected nodes
```

# Webscraping tutorial

```{r}
#install.packages("xml2")
require(xml2)
#install.packages("rvest")
require(rvest)
```
```{r}
url <- "https://nos.nl/zoeken?q=meertaligheid&page=1"

nos <- read_html(url)


#elements: alle elementen, element: één element
lijst <- nos %>% html_elements("main") %>%
  html_elements("ul") %>%
  html_elements("li")

lijst[1] %>% html_children %>% html_attr(name="href")

lijst[1] %>% html_elements("time") %>% 
  html_attr(name="datetime")

```

