---
title: "Week2"
author: "Susanna"
date: "2026-09-07"
output: html_document
---
## Setting up the Packages

```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
```

```{r}
#install.packages('RSiena')
library(RSiena)
#install.packages('igraph')
library(igraph)
```

## Playing with Descriptives

First I want to visualise the data because that makes it easier to understand:

```{r}
View(s501)
copys501 <- s501
graph1 <- graph_from_adjacency_matrix(
  copys501,
  mode = "undirected",
  weighted = NULL,
  diag = TRUE,
  add.colnames = NULL,
  add.rownames = NA
)
plot(graph1)
```

My mission was to explore the betweenness, so let's use the betweenness() function:

```{r}
betweenness(
  graph1,
  v = V(graph1),
  directed = FALSE,
  weights = NULL,
  normalized = FALSE,
  cutoff = -1
)
```
V22 has the highest betweenness score! 

Now let's try a directed version:

```{r}
graph2 <- graph_from_adjacency_matrix(
  copys501,
  mode = "directed",
  weighted = NULL,
  diag = TRUE,
  add.colnames = NULL,
  add.rownames = NA
)
betweenness(
  graph2,
  v = V(graph2),
  directed = TRUE,
  weights = NULL,
  normalized = FALSE,
  cutoff = -1
)
```
V22 still has the highest betweenness score, but the numbers are different now. I would love to know what this data actually means (what are the ties? Friends? Partners? Co-authors?) and whether the ties were measured as undirected or directed... What does R do when you tell it that your network has directed ties when really they are undirected?
